\section{Graph Theory}

\subsection{General properties}

\subsubsection{Degrees}

\begin{theorem} \label{sumIndegreesIsNrEdges}
    In a directed graph $\sum_{V_G} \indeg(v_n) = |E_G|$
\end{theorem}

\begin{proof}
    \induction{$ \sum_{V_G} \indeg(v_n) = |E_G| $ \\}
    {$\sum_{V_G} \indeg(v_n) = |E_G|$ where $E_G = \{(v_1, v_2)\}$ \\}
    {This is tirvially true.}
    {
    \begin{itemize}
        \item Let $\sum_{V_G} \indeg(v_n) = |E_G|$ 
        \item Let $V_{G'} = V_G \cup \{v_{N+1}\}$
        \item Let $E_{G'} = E_G \cup \{(x, y) \in V_G \times \{v_{N+1}\} \cup \{v_{N+1}\} \times V_G \}$
    \end{itemize}
    }
    {
     For the left side it holds that: 
                $$ \sum_{V_{G'}} \indeg(v_n) = \sum_{V_G} \indeg(v_n) + \outdeg(v_{N+1}) + \indeg(v_{N+1}) $$
                On the other hand on the right side it holds that: 
                $$ |E_{G'}| = |E_G| + \outdeg(v_{N+1}) + \indeg(v_{N+1}) $$
                So the equation reduces to $0=0$, which is true.
    }
\end{proof}


\begin{lemma} \label{indegIsOutdeg}
    $\sum_{V_G} \indeg(v_n) = \sum_{V_G} \outdeg(v_n)$
\end{lemma}

\begin{lemma}
    In an undirected graph we have $\sum_{V_G} \degr(v_n) = 2|E_G|$
\end{lemma}

\subsection{Walks and paths}

\begin{definition}
    A walk is any sequence of vertices that are connected by an edge.
\end{definition}

\begin{definition}
    A path is a walk where no vertex appears twice.
\end{definition}

\begin{theorem}
    The shortest walk between any two vertices in a graph is a path.
\end{theorem}

\begin{proof}
    \subprf{By contradiction. Suppose that the vertex $x$ appears twice in the shortest walk.}{this is a contradiction.}{
        Since $x$ appears twice, the proposed walk must be of the form $f-x-g-x-h$, where $g$ is a sequence of $0$ or more connected vertices. Then this walk can be shortened to $f-x-h$.
    }
\end{proof}

\begin{theorem}
    The longest path in a graph has length $ | V_G | - 1 $.
\end{theorem}
This is too trivial for a full-fledged proof: there are no repetitions in a path, so it can only have as many steps as it has vertices.


\subsubsection{Relations and adjacency matrices}

Let $R \circ Q$ be a composition of two relations. The number of paths of length exactly $n$ between two vertices $x$ and $y$, written as $\paths_n(x, z)$, can be expressed  as 
 $$ \paths_n(x,z) = | \{ y | xRy \land yQz \} | = \sum_{y \in Y} xRy \cdot yQz$$


If $R_{AM}$ and $Q_{AM}$ are the adjacency matrix representations of the above relations, then: 

$$  R_{AM} \cdot Q_{AM} [n,m] = \sum_{y \in Y} R_{AM} [x_n, y] \cdot Q_{AM} [y, z_m] $$

$$ = \sum_{y \in Y} xRy \cdot yQz = \paths_n(x,z) $$

When $R=Q$, it follows easily that:

\begin{theorem}
    $R_{AM}^2 [n,m]$ is the number of paths  that go from $v_n$ to $v_m$ in exactly $2$ steps.
\end{theorem}

\begin{theorem} \label{all_possible_paths}
    $R_{AM}$ explains whether or not two vertices $v$ and $u$ are connected. All possible paths are listed in $R^* = R_{AM} \cup R_{AM}^2 \cup R_{AM}^3 \cup ... \cup R_{AM}^{|V_G|-1}$.
    We can find the length of the shortest path between two vertices $u$ and $v$ by
\end{theorem}



\subsection{Planar graphs}

The following theorems all deal with planar, connected graphs, and build up to Eulers theorem. Well use $N$ for the number of nodes, $E$ for the number of edges, and $L$ for the number of loops.

\begin{theorem}
    Adding a edge to a planar, connected graph means adding one loop: $ \Delta N = 0 \land \Delta E = 1 \then \Delta L = 1$
\end{theorem}

\begin{theorem}
    Adding x edges means adding x loops: $ \Delta N = 0 \land \Delta E = x \then \Delta L = x $
\end{theorem}

\begin{proof}
    \subprf{}{$\Delta N = 0 \land \Delta E = x \then \Delta L = x$}{
        By induction on $\Delta E$. \\
        \subprf{Base case: $\Delta E = 1$}{$\Delta L = 1$}{
            By theorem 11.
        }
        \subprf{Let $\Delta N = 0 \land \Delta E = x \then \Delta L = x$.\\}
            {$\Delta N = 0 \land \Delta E = x+1 \then \Delta L = x+1$}{
            Consider the graph from the induction hypothesis, then apply theorem 12.
        }
    }
\end{proof}

\begin{theorem}
    When we add a new node to a graph and connect it to the graph using one or more edges, then the number of edges and loops behaves as such: $ \Delta N = 1 \then \Delta L = \Delta E - 1 $
\end{theorem}

\begin{proof}
    \subprf{}{$\Delta N = 1 \then \Delta L = \Delta E - 1$}{
        Let $ \Delta N = 1 $. \\
        By induction on $ \Delta E $. \\
        \subprf{Base case: $\Delta E = 1 $}{$ \Delta L = 0 $}{
            Graphically trivial.
        }
        \subprf{Induction step: $\Delta L = \Delta E - 1$}{$\Delta L' = \Delta E' - 1$}{
            $\Delta E ' = \Delta E + 1$ \\
            Using theorem 12, that means:
            $\Delta L' = \Delta L +1$ \\
            Thus: 
            $\Delta L + 1 = \Delta E + 1 - 1 $ \\
            Which reduces to: 
            $\Delta L = \Delta E - 1$, which is true by the induction hypothesis.
        }
    }
\end{proof}

\begin{theorem}
    Eulers theorem: $N + L = E + 1$
\end{theorem}

\begin{proof}
    \subprf{}{$N + L = E + 1$}{
        By induction on $N$. \\
        \subprf{Base case: $N=1$ or $N=2$}{$N + L = E + 1$}{
            If $N=1$: $1 + 0 = 0 + 1$ \\
            If $N=2$: $2 + 0 = 1 + 1$
        }
        \subprf{Induction step: Let $N + L = E + 1$.\\}{$N+1 + L' = E' + 1$}{
            Using theorem 13: $\Delta L' = \Delta E - 1$ \\
            Thus:
            $N + 1 + L + \Delta L = E + \Delta E + 1$ \\
            $N + L = E + 1$, which is true  by the induction hypothesis.
        }
    }
\end{proof}



\subsubsection{Trees}


Here is a fun little proof.

\begin{proof} In a tree, the mean number of children ($mcc$) is always equal to $1 - 1/N$.
    
    \subprf{}{$mcc = 1 - \frac{1}{N}$}{
        With what we have so far, this is almost trivial to prove. 
        The mean child count can be written as 
        $$ mcc = \frac{1}{N} \sum_N \outdeg{v_n} $$
        Using theorem \ref{sumIndegreesIsNrEdges} and lemma \ref{indegIsOutdeg}, this becomes
        $$ mcc = \frac{1}{N} E $$
        Eulers formula in a tree becomes $N = E + 1$, since there are no loops in trees. Using this: 
        $$ mcc = 1 - \frac{1}{N} $$
    }

\end{proof}

When training a binary classifier, we usually need a measure of defining which of several trees is a good representation of the data. The number of nodes, however, cannot be used as a measure, because it is always going to be $2n - 1$.
\begin{proof} In a binary tree, the number of leaves $n$ equals $\frac{N+1}{2}$, regardless of the architecture of the tree.

    \subprf{}{$n = \frac{N+1}{2}$}{
        We shall use the fact that in any tree, $\sum_{V_G} \indeg(v_n) = \sum_{V_G} \outdeg(v_n)$. \\

        Consider the leaves of the tree. They all have $\indeg(v_n) = 1$ and $\outdeg(v_n) = 0$. Thus $\sum_{\text{leaves}} \indeg(v_n) = n$ and $\sum_{\text{leaves}} \outdeg(v_n) = 0$. \\
        Now consider the nodes (there are $N-n$ of them). They all have $\outdeg(v_n) = 2$ and $\indeg(v_n) = 1$, except for the root-node. Thus $\sum_{\text{nodes}} \indeg(v_n) = N - n - 1$ and $\sum_{\text{nodes}} \outdeg(v_n) = 2(N - n)$. \\
        Equating the in- and out-degrees, we obtain:
        $$ n + N - n - 1 = 0 + 2(N - n) $$
        $$ n = \frac{N+1}{2} $$
    }

\end{proof}

