\section{Statistics}

\subsection{Correlation}
Assume an inner product space.

\subsection{Linear regression}

Assume that reality can me modelled by a model like this one: 

$$ \vec{y} = \mtrx{X} \vec{w} + \vec{e} $$

Not knowing $\vec{e}$, our best guess at the outcome would be 

$$ \vec{\hat{y}} = \mtrx{X} \vec{w} $$

We can find the gradient of w by:

$$ \partDiff{E}{\vec{w}} = -  (\vec{y} - \vec{\hat{y}}) \mtrx{X}  $$



\subsection{Principal component analysis}
Consider the case where you have a three-dimensional euclidean coordinate system, but all your observation-points form a plane inside this space. Try to find a new orthagonal base that minimizes the shadow of the datapoints.
It turns out that the best base you can find for this purpose is the set of eigenvectors of the datamatrix. 



\subsection{Tests}
Often, statistical tests work \emph{the wrong way round}. You have your hypothesis, like "these groups are different". Then you state the opposite, the \emph{null-hypothesis}. Then you determine how likely your data is under the null-hypothesis - the \emph{p-value}. And than you hope that the p-value is very low.

\paragraph{Students T-Test} compares two averages (means) and tells you if they are different from each other. The null hypothesis is that all samples come from the same distribution. That means that under $H_0$, the two means you obtained should be similar. So the interpretation goes like this: 
\begin{itemize}
    \item low p-value: the chance that your result would have happend under $H_0$ is low. It is very unlikely that the two means you have obtained actually come from the same distribution. 
    \item high p-value: It is quite possible that both your groups come from the same distribution. 
\end{itemize}
Note that there are two experimental setups that can use t-tests:
\begin{itemize}
    \item unpaired: you have two groups, with no person in both groups. Example: Test one group with medication and another with placebo. 
    \item paired: every person is first in the first and then in the second group. Example: test patients before and after treatment. 
\end{itemize}
